---
title: "kk"
author: "Avadhoot"
date: "6/15/2020"
output: html_document
editor_options: 
  chunk_output_type: console
---
```{r}
library(tidyverse)
library(stringr)
library(dplyr)
library(ggplot2)
library(ggrepel)
library(ggmap)
library(phenofit)
setwd("C:/Users/avdho/Desktop/SEM/DATA MINING/Project/data_to_work_with2")
table1 <- read.csv("H-1B_Disclosure_RAW_Data_FY15.csv",stringsAsFactors = FALSE)
table2 <- read.csv("H-1B_Disclosure_RAW_Data_FY16.csv",stringsAsFactors = FALSE)
table3 <- read.csv("H-1B_Disclosure_RAW_Data_FY17.csv",stringsAsFactors = FALSE)

table4 <- read.csv("H-1B_Disclosure_RAW_Data_FY18.csv",stringsAsFactors = FALSE)
getwd()

```

PREPARING DATASET
```{r}
names(table2)[24] <- 'TOTAL_WORKERS'
names(table2)[29] <- 'PW_SOURCE_YEAR'
names(table2)[30] <- 'PW_SOURCE_OTHER'
names(table2)[28] <- 'PW_SOURCE'
names(table2)[23] <- 'NAICS_CODE'
names(table2)[34] <- 'H1B_DEPENDENT'

```


```{r}
new <- table2[ ,!(colnames(table2) %in% c('CASE_NUMBER','CASE_STATUS','CASE_SUBMITTED','DECISION_DATE','VISA_CLASS','EMPLOYMENT_START_DATE','EMPLOYMENT_END_DATE','EMPLOYER_NAME','EMPLOYER_ADDRESS1','EMPLOYER_ADDRESS2','EMPLOYER_CITY','EMPLOYER_STATE','EMPLOYER_POSTAL_CODE','EMPLOYER_COUNTRY','EMPLOYER_PROVINCE','EMPLOYER_PHONE','EMPLOYER_PHONE_EXT','AGENT_ATTORNEY_NAME','AGENT_ATTORNEY_CITY','AGENT_ATTORNEY_STATE','JOB_TITLE','SOC_CODE','SOC_NAME','NAIC_CODE','TOTAL WORKERS','FULL_TIME_POSITION','PREVAILING_WAGE','PW_UNIT_OF_PAY','PW_WAGE_LEVEL','PW_WAGE_SOURCE','PW_WAGE_SOURCE_YEAR','PW_WAGE_SOURCE_OTHER','WAGE_RATE_OF_PAY','WAGE_UNIT_OF_PAY','H-1B_DEPENDENT','WILLFUL VIOLATOR','WORKSITE_CITY','WORKSITE_COUNTY','WORKSITE_STATE','WORKSITE_POSTAL_CODE'))]

common_cols <- intersect(colnames(table3),colnames(table4))
df <- rbind(table3[,common_cols],table4[, common_cols])

common_cols1 <- intersect(colnames(df),colnames(table2)) 
df_new <- rbind(df[,common_cols1],table2[, common_cols1])



colnames(table1)
colnames(df_new)
df_new <- df_new %>% filter( VISA_CLASS =="H-1B")
colnames(df_new)

df_new <- df_new[,c(-1,-5,-9,-12,-14,-15,-16,-17,-39)]
table1 <- table1[,c(-1,-9,-10,-13,-15,-16,-17,-18,-40)]
table1 <- table1 %>% filter( VISA_CLASS =="H-1B")
table1 <- table1[,c(-4)]
names(table1)[26]<-"H1B_DEPENDENT"
names(table1)[27]<-"WILLFUL_VIOLATOR"
names(table1)[15]<-"NAICS_CODE"
names(table1)[16] <- 'TOTAL_WORKERS'
names(table1)[22] <- 'PW_SOURCE_YEAR'
names(table1)[23] <- 'PW_SOURCE_OTHER'
names(table1)[21] <- 'PW_SOURCE'
unique(df_new$EMPLOYER_COUNTRY)    ##talk about employer country
table1<-cbind(as.data.frame(str_split_fixed(table1$WAGE_RATE_OF_PAY,"-",2)),table1)
colnames(table1)
names(table1)[1]<-"WAGE_RATE_OF_PAY_FROM"
names(table1)[2]<-"WAGE_RATE_OF_PAY_TO"

str(table1$WAGE_RATE_OF_PAY)

intersect_new <- intersect(colnames(df_new),colnames(table1))   #Join table1 with df_new--number of common coloumns
final <- rbind(df_new[,intersect_new],table1[, intersect_new])

```

```{r}
NROW(final$CASE_STATUS)

data.frame(colnames(final))
sapply(final, function(x) sum(is.na(x)))



# to replace missing values in columns by NA
#1) 
sum(final$CASE_SUBMITTED %in% "")
final$CASE_SUBMITTED <- replace(final$CASE_SUBMITTED, final$CASE_SUBMITTED == "", NA)
table(is.na(final$CASE_SUBMITTED))

#2)
sum(final$EMPLOYMENT_START_DATE %in% "")
final$EMPLOYMENT_START_DATE <- replace(final$EMPLOYMENT_START_DATE, final$EMPLOYMENT_START_DATE == "", NA)
table(is.na(final$EMPLOYMENT_START_DATE))

#3) 
sum(final$EMPLOYMENT_END_DATE %in% "")
final$EMPLOYMENT_END_DATE <- replace(final$EMPLOYMENT_END_DATE, final$EMPLOYMENT_END_DATE == "", NA)
table(is.na(final$EMPLOYMENT_END_DATE))

#4) 
sum(final$EMPLOYER_NAME %in% "")
final$EMPLOYER_NAME <- replace(final$EMPLOYER_NAME, final$EMPLOYER_NAME =="", NA)
table(is.na(final$EMPLOYER_NAME))

#5) 
sum(final$EMPLOYER_CITY %in% "")
final$EMPLOYER_CITY <- replace(final$EMPLOYER_CITY, final$EMPLOYER_CITY =="", NA)
table(is.na(final$EMPLOYER_CITY))

#6)
sum(final$EMPLOYER_STATE %in% "")
final$EMPLOYER_STATE <- replace(final$EMPLOYER_STATE, final$EMPLOYER_STATE == "", NA)
table(is.na(final$EMPLOYER_STATE))

#7) 
sum(final$EMPLOYER_COUNTRY %in% "")
final$EMPLOYER_COUNTRY <- replace(final$EMPLOYER_COUNTRY, final$EMPLOYER_COUNTRY =="", NA)
table(is.na(final$EMPLOYER_COUNTRY))

#8) 
unique(final$AGENT_ATTORNEY_CITY) # missing values -> ""
sum(final$AGENT_ATTORNEY_CITY %in% "")
final$AGENT_ATTORNEY_CITY <- replace(final$AGENT_ATTORNEY_CITY, final$AGENT_ATTORNEY_CITY == "", NA)
table(is.na(final$AGENT_ATTORNEY_CITY))

#9) 
sum(final$AGENT_ATTORNEY_STATE %in% "")
final$AGENT_ATTORNEY_STATE <- replace(final$AGENT_ATTORNEY_STATE, final$AGENT_ATTORNEY_STATE == "", NA)
table(is.na(final$AGENT_ATTORNEY_STATE))

#10)
sum(final$JOB_TITLE %in% "")
final$JOB_TITLE <- replace(final$JOB_TITLE, final$JOB_TITLE == "", NA)
table(is.na(final$JOB_TITLE))

#11)
sum(final$SOC_CODE %in% "")
final$SOC_CODE <- replace(final$SOC_CODE, final$SOC_CODE == "", NA)
table(is.na(final$SOC_CODE))

#12) 
sum(final$SOC_NAME %in% "")
final$SOC_NAME <- replace(final$SOC_NAME, final$SOC_NAME == "", NA)
table(is.na(final$SOC_NAME))

#13) 
sum(final$NAICS_CODE %in% "")
final$NAICS_CODE <- replace(final$NAICS_CODE, final$NAICS_CODE =="", NA)
table(is.na(final$NAICS_CODE))

#14)
sum(final$FULL_TIME_POSITION %in% "")
final$FULL_TIME_POSITION <- replace(final$FULL_TIME_POSITION, final$FULL_TIME_POSITION == "", NA)
table(is.na(final$FULL_TIME_POSITION))

#15)
sum(final$PREVAILING_WAGE %in% "")
final$PREVAILING_WAGE <- replace(final$PREVAILING_WAGE, final$PREVAILING_WAGE == "", NA)
table(is.na(final$PREVAILING_WAGE))

#16) 
sum(final$PW_UNIT_OF_PAY %in% "")
final$PW_UNIT_OF_PAY <- replace(final$PW_UNIT_OF_PAY, final$PW_UNIT_OF_PAY == "", NA)
table(is.na(final$PW_UNIT_OF_PAY))

#18)
sum(final$PW_SOURCE %in% "")
final$PW_SOURCE <- replace(final$PW_SOURCE, final$PW_SOURCE == "", NA)
table(is.na(final$PW_SOURCE))

#19) 
sum(final$PW_SOURCE_YEAR %in% "")
final$PW_SOURCE_YEAR <- replace(final$PW_SOURCE_YEAR, final$PW_SOURCE_YEAR == "", NA)
table(is.na(final$PW_SOURCE_YEAR))

#20) 
sum(final$PW_SOURCE_OTHER %in% "")
final$PW_SOURCE_OTHER <- replace(final$PW_SOURCE_OTHER, final$PW_SOURCE_OTHER == "", NA)
table(is.na(final$PW_SOURCE_OTHER))

#21) 
sum(final$WAGE_RATE_OF_PAY_FROM %in% "")
final$WAGE_RATE_OF_PAY_FROM <- replace(final$WAGE_RATE_OF_PAY_FROM, final$WAGE_RATE_OF_PAY_FROM == "", NA) 
table(is.na(final$WAGE_RATE_OF_PAY_FROM))

#22)
sum(final$WAGE_RATE_OF_PAY_TO %in% "")
final$WAGE_RATE_OF_PAY_TO <- replace(final$WAGE_RATE_OF_PAY_TO, final$WAGE_RATE_OF_PAY_TO == "", NA)
table(is.na(final$WAGE_RATE_OF_PAY_TO))

#23)
sum(final$WAGE_UNIT_OF_PAY %in% "")
final$WAGE_UNIT_OF_PAY <- replace(final$WAGE_UNIT_OF_PAY, final$WAGE_UNIT_OF_PAY == "", NA)
table(is.na(final$WAGE_UNIT_OF_PAY))

#24)
sum(final$H1B_DEPENDENT %in% "")
final$H1B_DEPENDENT <- replace(final$H1B_DEPENDENT, final$H1B_DEPENDENT == "", NA)
table(is.na(final$H1B_DEPENDENT))

#25) 
sum(final$WILLFUL_VIOLATOR %in% "")
final$WILLFUL_VIOLATOR <- replace(final$WILLFUL_VIOLATOR, final$WILLFUL_VIOLATOR =="", NA)
table(is.na(final$WILLFUL_VIOLATOR))

#26) 
sum(final$WORKSITE_CITY %in% "")
final$WORKSITE_CITY <- replace(final$WORKSITE_CITY, final$WORKSITE_CITY == "", NA)
table(is.na(final$WORKSITE_CITY))

#27) 
sum(final$WORKSITE_COUNTY %in% "")
final$WORKSITE_COUNTY <- replace(final$WORKSITE_COUNTY, final$WORKSITE_COUNTY == "", NA)
table(is.na(final$WORKSITE_COUNTY))

#28) 
sum(final$WORKSITE_STATE %in% "")
final$WORKSITE_STATE <- replace(final$WORKSITE_STATE, final$WORKSITE_STATE == "", NA)
table(is.na(final$WORKSITE_STATE))

str(final$AGENT_ATTORNEY_CITY)

final$CASE_SUBMITTED
final <- final %>% drop_na(-FULL_TIME_POSITION)
unique(final$CASE_STATUS)
kk<-final %>% filter(CASE_STATUS == "CERTIFIED"|CASE_STATUS =="DENIED")


```





```{r}

rmov_col <- kk[c(3,6,7,8,9,10,11,15,22,23,24,25,28,29)]
colnames(rmov_col)
kk <- kk[-c(3,6,7,8,9,10,11,15,22,23,24,25,28,29)]
data.frame(colnames(kk))
```


Removing all NAs
```{r}
str(kk)
sapply(kk, function(x) sum(is.na(x))) #gives NAs
summary(kk)

table(is.na(kk$FULL_TIME_POSITION)) #348136 NAs
unique(kk$FULL_TIME_POSITION)
summary(kk$FULL_TIME_POSITION)

kk %>%
  group_by(FULL_TIME_POSITION) %>%
  summarise(count = n(), percentage = 100*count/dim(kk)[1])
#there are total 31.3% NA values in FULL_TIME_POSITION
# lets see what happens when we remove all the rows with NAs
kk <- kk[!is.na(kk$FULL_TIME_POSITION),] #dataset without any NAs (removed all the NAs)
NROW(kk$CASE_STATUS)

summary(kk$FULL_TIME_POSITION)
summary(kk)
sapply(kk, function(x) sum(is.na(x)))
data.frame(colnames(kk))
# seems like we have a hell lot of rows to operate with...(762928 rows)

```


Creating Job Catagories
```{r}

kknew <- kk %>% mutate(NewSOCNAME = ifelse(str_detect(SOC_CODE,"11-"),'Management Occupations',SOC_NAME))
kknew <- kknew %>% mutate(NewSOCNAME = ifelse(str_detect(NewSOCNAME,"^11"),'Management Occupations',NewSOCNAME))
kknew <- kknew %>% mutate(NewSOCNAME = ifelse(str_detect(SOC_CODE,"^11"),'Management Occupations',NewSOCNAME))
kknew <- kknew %>% mutate(NewSOCNAME = ifelse(str_detect(SOC_CODE,"13-"),'Business and Financial Operations Occupations',NewSOCNAME))
kknew <- kknew %>% mutate(NewSOCNAME = ifelse(str_detect(NewSOCNAME,"^13"),'Business and Financial Operations Occupations',NewSOCNAME))
kknew <- kknew %>% mutate(NewSOCNAME = ifelse(str_detect(SOC_CODE,"^13"),'Business and Financial Operations Occupations',NewSOCNAME))
kknew <- kknew %>% mutate(NewSOCNAME = ifelse(str_detect(SOC_CODE,"15-"),'Computer and Mathematical Operations Occupations',NewSOCNAME))
kknew <- kknew %>% mutate(NewSOCNAME = ifelse(str_detect(NewSOCNAME,"^15"),'Computer and Mathematical Operations Occupations',NewSOCNAME))
kknew <- kknew %>% mutate(NewSOCNAME = ifelse(str_detect(SOC_CODE,"^15"),'Computer and Mathematical Operations Occupations',NewSOCNAME))

kknew <- kknew %>% mutate(NewSOCNAME = ifelse(str_detect(SOC_CODE,"17-"),'Architecture and Engineering Occupations',NewSOCNAME))
kknew <- kknew %>% mutate(NewSOCNAME = ifelse(str_detect(NewSOCNAME,"^17"),'Architecture and Engineering Occupations',NewSOCNAME))
kknew <- kknew %>% mutate(NewSOCNAME = ifelse(str_detect(SOC_CODE,"^17"),'Architecture and Engineering Occupations',NewSOCNAME))

kknew <- kknew %>% mutate(NewSOCNAME = ifelse(str_detect(SOC_CODE,"19-"),'Life,Physical,and Social Science Occupations',NewSOCNAME))
kknew <- kknew %>% mutate(NewSOCNAME = ifelse(str_detect(NewSOCNAME,"^19"),'Life,Physical,and Social Science Occupations',NewSOCNAME))

kknew <- kknew %>% mutate(NewSOCNAME = ifelse(str_detect(SOC_CODE,"21-"),'Community and Social Service Occupations',NewSOCNAME))
kknew <- kknew %>% mutate(NewSOCNAME = ifelse(str_detect(NewSOCNAME,"^21"),'Community and Social Service Occupations',NewSOCNAME))

kknew <- kknew %>% mutate(NewSOCNAME = ifelse(str_detect(NewSOCNAME,"^23"),'Legal Occupations',NewSOCNAME))

kknew <- kknew %>% mutate(NewSOCNAME = ifelse(str_detect(SOC_CODE,"23-"),'Legal Occupations',NewSOCNAME))
kknew <- kknew %>% mutate(NewSOCNAME = ifelse(str_detect(SOC_CODE,"23-"),'Legal Occupations',NewSOCNAME))
kknew <- kknew %>% mutate(NewSOCNAME = ifelse(str_detect(NewSOCNAME,"15-1121"),'Computer and Mathematical Operations Occupations',NewSOCNAME))
kknew <- kknew %>% mutate(NewSOCNAME = ifelse(str_detect(SOC_CODE,"25-"),'Educational Instrction and Library Occupations',NewSOCNAME))

kknew <- kknew %>% mutate(NewSOCNAME = ifelse(str_detect(SOC_CODE,"27-"),'Arts, Design,Entertainment,Sports, and Media Occupations',NewSOCNAME))
kknew <- kknew %>% mutate(NewSOCNAME = ifelse(str_detect(NewSOCNAME,"^27"),'Arts, Design,Entertainment,Sports, and Media Occupations',NewSOCNAME))

kknew <- kknew %>% mutate(NewSOCNAME = ifelse(str_detect(SOC_CODE,"29-"),'Healthcare Practitioners and Technical Occupations',NewSOCNAME))
kknew <- kknew %>% mutate(NewSOCNAME = ifelse(str_detect(NewSOCNAME,"^29"),'Healthcare Practitioners and Technical Occupations',NewSOCNAME))
kknew <- kknew %>% mutate(NewSOCNAME = ifelse(str_detect(SOC_CODE,"^29"),'Healthcare Practitioners and Technical Occupations',NewSOCNAME))

kknew <- kknew %>% mutate(NewSOCNAME = ifelse(str_detect(SOC_CODE,"31-"),'Healthcare Support Occupations',NewSOCNAME))
kknew <- kknew %>% mutate(NewSOCNAME = ifelse(str_detect(NewSOCNAME,"^31"),'Healthcare Support Occupations',NewSOCNAME))

kknew <- kknew %>% mutate(NewSOCNAME = ifelse(str_detect(SOC_CODE,"33-"),'Protective Service Occupations',NewSOCNAME))
kknew <- kknew %>% mutate(NewSOCNAME = ifelse(str_detect(NewSOCNAME,"^33"),'Protective Service Occupations',NewSOCNAME))

kknew <- kknew %>% mutate(NewSOCNAME = ifelse(str_detect(SOC_CODE,"35-"),'Food Preparation and Serving Related Occupations',NewSOCNAME))
kknew <- kknew %>% mutate(NewSOCNAME = ifelse(str_detect(NewSOCNAME,"^35"),'Food Preparation and Serving Related Occupations',NewSOCNAME))


kknew <- kknew %>% mutate(NewSOCNAME = ifelse(str_detect(SOC_CODE,"37-"),'Building and Grounds Cleaning Maintenance Occupations',NewSOCNAME))
kknew <- kknew %>% mutate(NewSOCNAME = ifelse(str_detect(NewSOCNAME,"^37"),'Building and Grounds Cleaning Maintenance Occupations',NewSOCNAME))

kknew <- kknew %>% mutate(NewSOCNAME = ifelse(str_detect(SOC_CODE,"39-"),'Persoanl Care and Service Occupations',NewSOCNAME))
kknew <- kknew %>% mutate(NewSOCNAME = ifelse(str_detect(NewSOCNAME,"^39"),'Persoanl Care and Service Occupations',NewSOCNAME))

kknew <- kknew %>% mutate(NewSOCNAME = ifelse(str_detect(SOC_CODE,"41-"),'Sales and Related Occupations',NewSOCNAME))
kknew <- kknew %>% mutate(NewSOCNAME = ifelse(str_detect(NewSOCNAME,"^41"),'Sales and Related Occupations',NewSOCNAME))

kknew <- kknew %>% mutate(NewSOCNAME = ifelse(str_detect(SOC_CODE,"43-"),'Office and Administrative Support Occupations',NewSOCNAME))
kknew <- kknew %>% mutate(NewSOCNAME = ifelse(str_detect(NewSOCNAME,"^43"),'Office and Administrative Support Occupations',NewSOCNAME))
kknew <- kknew %>% mutate(NewSOCNAME = ifelse(str_detect(NewSOCNAME,"COMPUTER"),'Computer and Mathematical Operations Occupations',NewSOCNAME))


kknew <- kknew %>% mutate(NewSOCNAME = ifelse(str_detect(SOC_CODE,"45-"),'Farming Fishing and Forestry Occupations',NewSOCNAME))
kknew <- kknew %>% mutate(NewSOCNAME = ifelse(str_detect(NewSOCNAME,"^45"),'Farming Fishing and Forestry Occupations',NewSOCNAME))

kknew <- kknew %>% mutate(NewSOCNAME = ifelse(str_detect(SOC_CODE,"47-"),'Construction and Extraction Occupations',NewSOCNAME))
kknew <- kknew %>% mutate(NewSOCNAME = ifelse(str_detect(NewSOCNAME,"^47"),'Construction and Extraction Occupations',NewSOCNAME))

kknew <- kknew %>% mutate(NewSOCNAME = ifelse(str_detect(SOC_CODE,"49-"),'Installation, Repair and Maintenance Occupations',NewSOCNAME))
kknew <- kknew %>% mutate(NewSOCNAME = ifelse(str_detect(NewSOCNAME,"^49"),'Installation, Repair and Maintenance Occupations',NewSOCNAME))

kknew <- kknew %>% mutate(NewSOCNAME = ifelse(str_detect(SOC_CODE,"51-"),'Production Occupations',NewSOCNAME))
kknew <- kknew %>% mutate(NewSOCNAME = ifelse(str_detect(NewSOCNAME,"^51"),'Production Occupations',NewSOCNAME))

kknew <- kknew %>% mutate(NewSOCNAME = ifelse(str_detect(SOC_CODE,"53-"),'Transportation and Material Moving Occupations',NewSOCNAME))
kknew <- kknew %>% mutate(NewSOCNAME = ifelse(str_detect(NewSOCNAME,"^53"),'Transportation and Material Moving Occupations',NewSOCNAME))
kknew <- kknew %>% mutate(NewSOCNAME = ifelse(!str_detect(SOC_CODE,"-"),'NA',NewSOCNAME))

kknew <- kknew %>% mutate(NewSOCNAME = ifelse(str_detect(NewSOCNAME,"SOFTWARE"),'Production Occupations',NewSOCNAME))

kknew <- kknew %>% mutate(NewSOCNAME = ifelse(str_detect(NewSOCNAME,"ENGINEERS"),'Architecture and Engineering Occupations',NewSOCNAME))

kknew <- kknew %>% mutate(NewSOCNAME = ifelse(str_detect(NewSOCNAME,"SOFTWARE"),'Computer and Mathematical Operations Occupations',NewSOCNAME))
kknew <- kknew %>% mutate(NewSOCNAME = ifelse(str_detect(NewSOCNAME,"ENGINEERS"),'Architecture and Engineering Occupations',NewSOCNAME))
kknew <- kknew %>% mutate(NewSOCNAME = ifelse(str_detect(NewSOCNAME,"PEDIATRICIANS"),'Healthcare Practitioners and Technical Occupations',NewSOCNAME))
kknew <- kknew %>% mutate(NewSOCNAME = ifelse(str_detect(NewSOCNAME,"PRACTITIONERS"),'Healthcare Practitioners and Technical Occupations',NewSOCNAME))
kknew <- kknew %>% mutate(NewSOCNAME = ifelse(str_detect(NewSOCNAME,"SURGEONS"),'Healthcare Practitioners and Technical Occupations',NewSOCNAME))
kknew<-kknew %>% filter(NewSOCNAME!="NA")

kknew[kknew$SOC_CODE=="",]<-NA
kknew<-drop_na(kknew)

unique(kknew$NewSOCNAME)

h1b <- kknew

rm(table1, table2, table3, table4)

```




```{r}

#CONVERSION OF DATES TO NUMERIC VALUES
h1b$CASE_SUBMITTED <- as.POSIXct(h1b$CASE_SUBMITTED,format = "%m/%d/%Y")
#h1b$DECISION_DATE <- as.POSIXct(h1b$DECISION_DATE,format = "%m/%d/%Y")

h1b$EMPLOYMENT_START_DATE <- as.POSIXct(h1b$EMPLOYMENT_START_DATE,format = "%m/%d/%Y")
h1b$EMPLOYMENT_END_DATE <- as.POSIXct(h1b$EMPLOYMENT_END_DATE,format = "%m/%d/%Y")

#converting to numeric data type
#h1b$WAGE_RATE_OF_PAY_FROM<-as.numeric(as.factor(h1b$WAGE_RATE_OF_PAY_FROM))
#h1b$WAGE_RATE_OF_PAY_TO<-as.numeric(as.factor(h1b$WAGE_RATE_OF_PAY_TO))
#h1b$PREVAILING_WAGE<-as.numeric(as.factor(h1b$WAGE_RATE_OF_PAY_TO))
h1b$TOTAL_WORKERS<-as.numeric(as.factor(h1b$TOTAL_WORKERS))
h1b$PW_SOURCE_YEAR<-as.numeric(as.factor(h1b$PW_SOURCE_YEAR))

```


Create new variables:
SUBMITTED_MONTH = month on which the case was submitted
Employment_time = EMPLOYMENT_END_DATE - EMPLOYMENT_START_DATE

```{r}
h1b %>%
  group_by(CASE_SUBMITTED) %>%
  summarise(count = n()) %>%
  arrange(desc(count))

str(h1b)
#h1b$process_time <- as.Date(as.character(h1b$DECISION_DATE), format="%Y-%m-%d") -
#  as.Date(as.character(h1b$CASE_SUBMITTED), format = "%Y-%m-%d")

#unique(h1b$process_time)                               

library (lubridate)
#CONVERSION OF DATES TO NUMERIC VALUES
h1b$CASE_SUBMITTED <- as.POSIXct(h1b$CASE_SUBMITTED,format = "%m/%d/%Y")
h1b$YEAR <- year(h1b$CASE_SUBMITTED)
h1b$SUBMITTED_MONTH <- month(h1b$CASE_SUBMITTED)


h1b$Employment_Time <- as.Date(as.character(h1b$EMPLOYMENT_END_DATE), format="%Y-%m-%d") - as.Date(as.character(h1b$EMPLOYMENT_START_DATE), format = "%Y-%m-%d")

max(h1b$Employment_Time)

data.frame(colnames(h1b))

h1b <- h1b[-c(2,3,4,5,6,7,11,13,18)] #removing YEAR for model exp
data.frame(colnames(h1b))

```


```{r}
#rm(kk, kknew, new, rmov_col,df, df_new, final)

#LOGISTIC REGRESSION :
data.frame(colnames(h1b))
library(ggplot2)
library(cowplot)
# to check how many values each observation has for CERTIFIED / DENIED condition 
xtabs(~CASE_STATUS + FULL_TIME_POSITION, data = h1b)
xtabs(~CASE_STATUS + TOTAL_WORKERS, data = h1b)
xtabs(~CASE_STATUS + PW_SOURCE, data = h1b)
xtabs(~CASE_STATUS + H1B_DEPENDENT, data = h1b)
xtabs(~CASE_STATUS + WILLFUL_VIOLATOR, data = h1b)
xtabs(~CASE_STATUS + NewSOCNAME, data = h1b)
xtabs(~CASE_STATUS + SUBMITTED_MONTH, data = h1b)

# now lets make test and train sets
h1b %>%
  group_by(CASE_STATUS) %>%
  summarise(count = n()) %>%
  arrange(desc(count))
data.frame(colnames(h1b))
h1b <- h1b[-c(8)]

# there are only 9643 DENIED cases in the set. Dataset is highly imbalanced 
# we need to balance the dataset while we split it 
set.seed(123)
# now we need to conver the variables into numerical 
library(caret)
l <- h1b[c(1,3,5,6,7,8,9)]
m <- h1b[-c(1,3,5,6,7,8,9)]

dmy <- dummyVars("~.", data = l, fullRank = T)
trsf <- data.frame(predict(dmy, newdata = l))
trsf

data.frame(colnames(m))
logist_reg <- cbind(trsf, m) # the full dataset with dummies
data.frame(colnames(logist_reg))

#rm(dmy, h1b, l, m, trsf) # removing all datasets which i will not use here


###
library(DMwR)
library(caret)
str(logist_reg)
prop.table(table(logist_reg$CASE_STATUSDENIED))


set.seed(1234)
splitIndex <- createDataPartition(logist_reg$CASE_STATUSDENIED, p=0.6, 
                                  list=FALSE, 
                                  times = 1)
trainSplit <- logist_reg[splitIndex,]
testSplit <- logist_reg[-splitIndex,]

table(trainSplit$CASE_STATUSDENIED)
prop.table(table(trainSplit$CASE_STATUSDENIED))

```



```{r}
###ROSE:
library(ROSE)
table(trainSplit$CASE_STATUSDENIED)
NROW(logist_reg)
NROW(trainSplit)

# BOTH:
#train_both <- ovun.sample(CASE_STATUSDENIED~., data = trainSplit, method = "both")$data
#table(train_both$target)
#prop.table(table(train_both$CASE_STATUSDENIED))

#trying to apply logistic regression to this training dataset
#train_both$target <- factor(train_both$CASE_STATUSDENIED)
#train_both <- train_both[-c(1)]
#model1 <- glm(target~., data = train_both, family = "binomial") #dooes not run cause its a big dataset


# 

### UNDERSAMPLING:

train_under <- ovun.sample(CASE_STATUSDENIED~., data = trainSplit, method = "under")$data
table(train_under$CASE_STATUSDENIED)

model2 <- glm(CASE_STATUSDENIED~., data = train_under, family = "binomial")

```
















VISUALIZATIONS
```{r}
h1b %>% ggplot(aes(x = CASE_STATUS)) +
  geom_bar()
#As we can see here,our data is highly imbalanced like hell, so we have to balance it first

#applications per year
h1b %>% ggplot(aes(x = YEAR)) +
  geom_bar() 

h1b %>% ggplot(aes(x = FULL_TIME_POSITION)) +
  geom_bar() 


h1b %>% ggplot(aes(y = WORKSITE_STATE)) +
  geom_bar()

```












Creating dummy variables
```{r}
data.frame(colnames(kknew))
library(caret)
#FULL_TIME, PW_SOURCE, H1B_DEPENDENT, WWLLFUL_VIOLATOR, [NewSOCNAMES]

l <- kknew[c(8,10,12,13,16)]
m <- kknew[-c(8,10,12,13,16)]
#
dmy <- dummyVars("~.", data = l, fullRank = T)
trsf <- data.frame(predict(dmy, newdata = l))
trsf
#

data.frame(colnames(m))
kk2 <- cbind(trsf, m)
data.frame(colnames(kk2))
kk2 <- kk3
kk3 <- kk3[-c(33,34,37)]
```


SPLIT DATASET
# at this stage out data is ready.....Almost!!
# We need to split the data into train(60%) and test(40%) datasets

```{r}
set.seed(123)
n <- nrow(kk3)
shuffled_df <- kk3[sample(n), ]
train_indices <- 1:round(0.6 * n)
train <- shuffled_df[train_indices, ]
test_indices <- (round(0.6 * n) + 1):n
test <- shuffled_df[test_indices, ]

NROW(kk3$CASE_STATUS)
NROW(train$CASE_STATUS)
NROW(test$CASE_STATUS)

```

LET'S VISUALIZE
```{r}
train %>% ggplot(aes(x = CASE_STATUS)) +
  geom_bar()

train %>%
  group_by(CASE_STATUS) %>%
  summarise(count = n(), percentage = 100*count/dim(kk)[1]) %>%
  arrange(desc(count))

```
As we can see here,our data is highly imbalanced like hell, so we have to balance it first
